\section{Motivation}

% - passwords are really important
% - everyone uses them
% - obvious upside: sharing

The most common way for humans to authenticate to digital systems nowadays is
the use of passwords, the core reason for which is its obvious simplicity.
Compared to other methods such as cryptographic keys, devices, or biometric
feature detection, it is very \emph{simple to understand} how passwords work
and how to use them, at least on a basic, rather insecure level.

Furthermore, passwords can be shared with other people. While this practice is
generally discouraged for security reasons, it still finds widespread use
within teams. Instead of creating separate accounts for each team member, and
updating them whenever someone joins or leaves the team, a group password is
generated and everyone in the team is given access via this password. This
reduces the management overhead for the team leaders, and it becomes very
important when different organizations work together using resources made
available through one of the organization's digital services.

% - done everywhere (team wide passwords)
% - obvious downside: security
% - solution: complicated + password manager

A problem arises when one aims at using passwords in a secure manner.  Given
the ever-increasing computational capacity of computers, one should not utilize
simple passwords that are easy to remember for authentication. Instead, it is
considered best practice to generate long random passwords from a large
character set, and not passwords them for different systems. This ensures that
the password can neither be guessed nor brute-forced in reasonable amounts of
time. To achieve this, many companies enforce a strict password guideline among
their employees. Because it is near impossible to remember a large number of
those high-quality passwords, password managers have come into existence, which
store all of a user's passwords using encryption technologies. Then, only a
single \emph{master password} is required to unlock all personal passwords
stored inside, which the user must remember (and keep safe).

% - sharing capability required
We consider a team of employees, each utilizing a password manager (as enforced
by company policy) to manage their personal passwords. Sharing team-wide
passwords via (possibly unencrypted) email or other text-based media is always
a security risk, and sharing them via phone or in person is tedious and may
result in errorneous transmission. Instead, the company seeks to use a common
password manager that is capable of sharing passwords with other people.

% - existing solutions: proprietary, cloud based
There are numerous solutions for this, which allow transmission of a password
secretly to a single contact, or group of contacts. However, these existing
solutions have one problem in common: they are all proprietary and often paid
solutions, hosted by their creators \emph{,,in the cloud''}, that is, on a
publicly accessibly system. While this makes it very simple to set up a team to
use these solutions, and generally works pretty well for everyone, it also
involves a security risk: highly confidential data, the secrets used for
authentication, are stored on foreign servers, made accessible via the
internet, outside the control of the data owners.

% - problem with those: highly confidential data not under company control
Furthermore, these servers are an interesting target for attackers, since they
store many secrets for large companies. The popular password management site
LastPass has just recently (in 2015) been target of an attack, where login
hashes have been exposed\cite{lastpass-security-notice}. Like that, a single
successful attack might allow access to many more protected resources --
LastPass claims to serve ,,millions of users''.

% - solution: distributed system
In my thesis, I plan to design a system that does not involve this security
risk: a distributed secret management system with sharing capabilities.

